---
title: "classification"
author: "Timothy Roy"
date: '2023-04-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Libraries
library(skimr)
```

# Task 1.1 Data cleaning 

## The first problem with the data is that the columns are not seperated 
```{r }
contract <- read.csv("labor-negotiations_students.csv")
head(contract)
```
## By using ';' seperator, the values can be organized into different columns

```{r}
contract <- read.csv("labor-negotiations_students.csv",  sep = ";")
head(contract)
str(contract)
```
## Removing NA values
```{r}
test <- read.csv("labor-negotiations_students.csv",sep = ";")

test[complete.cases(test),]
str(test)
```


## Checking for unique values to find anomalies

```{r}
num_of_cols <- ncol(contract)
print(paste("The number of columns in the df is ",num_of_cols ))


```

## Possible problem with 501?
* 501 stands out here because. It has probably replaced 16 here
* And 501 is missing after 500. 
*Therefore, I will use my intuition and replace them 
```{r}
unique(contract$profession)
```
## Possible fix?

```{r}
contract$profession <- seq(1,600, by = 1)

head(contract$profession, 20)
contract[500:520,]$profession
```


```{r}
unique(contract$duration)
```

## Removing NA values
```{r}
contract %>% skim()
```


