---
title: "classification"
author: "Timothy Roy"
date: '2023-04-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Libraries
library(skimr)
library(dplyr)
library(utils)
library(caret)
library(ggplot2)
library(caret)
library(lattice)
library(Matrix)
library(tidyr)
library(rpart)
```

# Task 1.1 Data cleaning 

## The first problem with the data is that the columns are not seperated 
```{r }
contract <- read.csv("labor-negotiations_students.csv")

head(contract)
```
## By using ';' seperator, the values can be organized into different columns

```{r}
contract <- read.csv("labor-negotiations_students.csv",  sep = ";")

head(contract)
str(contract)
```
```{r}
contract_backup<- read.csv("labor-negotiations_students.csv",  sep = ";")
str(contract_backup)
```

## Data Quality Issues
* Not removing rows if a columns have NA values since we have a small data set
* When using columns with NA values, all the NA values would be ignored
* Wage1 has negative values (2.0-7.0)
* Wage2 has value bigger than 7.0
* Wage3 has negative value
* Hours shouldn't be more than 40
* Stable pay consists of negative values
* Holidays should not contain more than 15 days


```{r}
contract %>% skim()
```

## Possible problem with 501?
* 501 stands out here because. It has probably replaced 16 here
* And 501 is missing after 500. 
*Therefore, I will use my intuition and replace them 

```{r}
contract$profession <- seq(1,600, by = 1)

head(contract$profession, 20)
contract[500:520,]$profession

```


## Removing columns with too many missing values

```{r}
contract <- subset(contract, select = -c(wage2,wage3,stby_pay,shift_diff, cola,pension,educ_allow, lngtrm_disabil,dntl_ins,breavement,Empl.hplan ))

contract
```


```{r}
contract <- na.omit(contract)
str(contract)
```



## Checking for unique values to find anomalies

```{r}
num_of_cols <- ncol(contract)
print(paste("The number of columns in the df is ",num_of_cols ))


```


```{r}
unique(contract$duration)
```

## Problem: Negative Values
* The description page shows that this columns can not contain negative values. Wage inc can't be negatiev. Therefore assuming it was a mistake, the negative values would be converted to positive values

```{r}
unique(contract$wage1)
max(contract$wage1)
min(abs(contract$wage1))
```
```{r}
contract$wage1 <- abs(contract$wage1)
contract$wage1 <- ifelse(contract$wage1 > 1.5 & contract$wage1 < 2, 2, contract$wage1)
contract$wage1 <- ifelse(contract$wage1 > 7 & contract$wage1 < 8, 7, contract$wage1)
contract$wage1
```


## Problem: hours more than 40 hours should not be here
```{r}
table(contract$hours)
#contract$hours
```

* Replacing values more than 40 and less than 30 with NA

```{r}

contract$hours <- abs(contract$hours)
contract$hours <- ifelse(contract$hours > 25 & contract$hours < 35, 35, contract$hours)
contract$hours <- ifelse(contract$hours > 40 & contract$hours < 50, 40, contract$hours)


contract$hours <- replace(contract$hours, contract$hours > 40, NA)
contract$hours <- replace(contract$hours, contract$hours < 35, NA)

table(contract$hours)
unique(contract$hours)
```

```{r}
contract <- na.omit(contract)
str(contract)
```

## Problem : out of bounds range
```{r}
unique(contract$holidays)
```

```{r}
contract$holidays <- abs(contract$holidays)
contract$holidays <- ifelse(contract$holidays > 5 & contract$holidays < 9, 9, contract$holidays)
contract$holidays <- ifelse(contract$holidays > 15 & contract$holidays < 22, 15, contract$holidays)


contract$holidays <- replace(contract$holidays, contract$holidays > 15, NA)
contract$holidays <- replace(contract$holidays, contract$holidays < 9, NA)

unique(contract$holidays)
```

```{r}
contract <- na.omit(contract)
str(contract)
```


## Problems: values are not in the correct forms
```{r}
unique(contract$vacation)
```

```{r}
contract$vacation <- replace(contract$vacation, contract$vacation == "below average", "ba")
contract$vacation <- replace(contract$vacation, contract$vacation == "average", "avg")
contract$vacation <- replace(contract$vacation, contract$vacation == "average", "gnr")
contract$vacation <- na_if(contract$vacation, "")
unique(contract$vacation)
```

```{r}
contract <- na.omit(contract)
str(contract)
```



```{r}
unique(contract$consent)
```


```{r}
contract$consent <- replace(contract$consent, contract$consent == "??", NA)

unique(contract$consent)

```

```{r}
contract <- na.omit(contract)
str(contract)
```
## Look at the data one more time :)

* All the out of ranges values have been removed
* Standard deviation has decreased for hours, stable, pay and more!

```{r}
contract %>% skim()
```



```{r}
contract
```

# Task 1.2 Classification


```{r}
# Convert the customer ID and contract status to factors

contract$profession <- as.factor(contract$profession )
contract$consent <- as.factor(contract$consent)
# Split the data into training and testing sets
set.seed(123)
train_index <- sample(nrow(contract), 0.8 * nrow(contract))
train_data <- contract[train_index, ]

test_data <- contract[-train_index, ]

#contract[train_index, ]
#contract[-train_index, ]




```

```{r}
train_data
```

```{r}
model <- rpart(consent ~ ., data = train_data, method = "class")
predictions <- predict(model, newdata = test_data, type = "class")
# Evaluate the model performance
confusionMatrix(predictions, test_data$consent)
```
```{r}
predictions_num <- as.numeric(predictions)

# Convert numeric predictions to "good" or "bad" based on threshold of 0.5
predictions_threshold <- ifelse(predictions_num > 0.5, "good", "bad")

# Evaluate the model performance
confusionMatrix(predictions_threshold, test_data$consent)
```



```{r}
set.seed(123)
train_index <- sample(nrow(contract), 0.8 * nrow(contract))
train_data <- contract[train_index, ]
test_data <- contract[-train_index, ]

# Ensure all factor variables have the same levels in both datasets
for (col in names(train_data)) {
  if (is.factor(train_data[[col]]) && is.factor(test_data[[col]])) {
    levels_test <- levels(test_data[[col]])
    levels_train <- levels(train_data[[col]])
    if (!identical(levels_test, levels_train)) {
      all_levels <- union(levels_test, levels_train)
      train_data[[col]] <- factor(train_data[[col]], levels = all_levels)
      test_data[[col]] <- factor(test_data[[col]], levels = all_levels)
    }
  }
}

model <- rpart(consent ~ ., data = train_data, method = "class")
predictions <- predict(model, newdata = test_data, type = "class")

# Convert predictions to factors with the same levels as test_data$consent
predictions <- factor(predictions, levels = levels(test_data$consent))

# Evaluate the model performance
confusionMatrix(predictions, test_data$consent)


```


```{r}
# Build a logistic regression model to predict the contract status
model <- glm(consent ~ ., data = train_data, family = "binomial")
```

```{r}

 
# Make predictions on the test data
predictions <- predict(model, newdata = test_data, type = "response")
predictions <- ifelse(predictions > 0.5, "good", "bad")
 
# Evaluate the model performance
confusionMatrix(predictions, test_data$consent)
```















